---
title: "NFL Melbourne 2026: reproducible analysis"
format:
  html:
    code-fold: false
    toc: true
editor_options:
  chunk_output_type: console
---

# Setup

This notebook explains how the study’s game data were used to calculate the results reported in the manuscript. It begins with game selection and the measures of performance, then describes the comparisons with opponents and season averages. Results are presented first for Melbourne and then for the historical London games.

Schedule, scoring, closing point-spread, and play-by-play data were obtained from nflverse on September 24, 2026. The analysis uses the saved copies from that date. Data from the 2010, 2013, 2017, 2019, and 2025 regular seasons provide the season comparisons and information about opponents. The 2026 data provide the Melbourne game and each team’s first following game.

The Melbourne game brought together two teams with different arrival schedules: the 49ers arrived early and spent the week in Melbourne, whereas the Rams arrived shortly before the game. Selected London games provide historical context for these approaches. The analysis includes the 49ers’ 2010 and 2013 games and the Rams’ 2017, 2019, and 2025 games, when each team used an arrival schedule similar to its Melbourne approach.

The 49ers’ Mexico City games were excluded because they involved shorter journeys and less time-zone displacement. The Rams’ 2012 and 2016 London games were excluded because the team arrived early on those trips.

The first regular-season game after each trip is also included. Together, these games provide 14 team performances: ten from the historical episodes and four from Melbourne and the two following games. Melbourne contributes two performances because both teams are included in the study.

```{r}
#| label: setup
#| code-fold: true

suppressPackageStartupMessages({
  library(tidyverse)
  library(patchwork)
})
```

```{r}
#| label: file-utilities
#| purl: true
#| code-fold: true

# checks that project is open and allows user to run notebook directly without changing paths
project_path <- function(...) {
  root <- if (file.exists("_quarto.yml")) "." else ".."

  if (!file.exists(file.path(root, "_quarto.yml"))) {
    stop("Open nfl-melbourne-2026-travel-performance.Rproj before running the analysis.")
  }

  return(file.path(normalizePath(root), ...))
}

# once path is correct we can load helpers
source(project_path("R", "figure_helpers.R"))
source(project_path("R", "historical_figures.R"))
source(project_path("R", "melbourne_performance_figure.R"))
source(project_path("R", "contextual_tables.R"))
```

```{r}
df_game_registry <- tribble(
  ~season, ~team, ~phase, ~game_id, ~expected_opponent,
  ~expected_week, ~expected_points_for, ~expected_points_against,
  2010, "SF", "international", "2010_08_DEN_SF", "DEN", 8,  24, 16,
  2010, "SF", "return",        "2010_10_STL_SF", "LA",  10, 23, 20,
  2013, "SF", "international", "2013_08_SF_JAX", "JAX", 8,  42, 10,
  2013, "SF", "return",        "2013_10_CAR_SF", "CAR", 10, 9,  10,
  2017, "LA", "international", "2017_07_ARI_LA", "ARI", 7,  33,  0,
  2017, "LA", "return",        "2017_09_LA_NYG", "NYG", 9,  51, 17,
  2019, "LA", "international", "2019_08_CIN_LA", "CIN", 8,  24, 10,
  2019, "LA", "return",        "2019_10_LA_PIT", "PIT", 10, 12, 17,
  2025, "LA", "international", "2025_07_LA_JAX", "JAX", 7,  35, 7,
  2025, "LA", "return",        "2025_09_NO_LA",  "NO",  9,  34, 10,
  2026, "SF", "international", "2026_01_SF_LA",  "LA",  1,  27, 7,
  2026, "LA", "international", "2026_01_SF_LA",  "SF",  1,  7,  27,
  2026, "SF", "return",        "2026_02_MIA_SF", "MIA", 2,  35, 13,
  2026, "LA", "return",        "2026_02_NYG_LA", "NYG", 2,  28, 6
)

df_reference_team_seasons <- tribble(
  ~season, ~team,
  2010, "SF",
  2013, "SF",
  2017, "LA",
  2019, "LA",
  2025, "LA",
  2025, "SF"
)
```

```{r}
#| label: archived-inputs
#| code-fold: true

snapshot_date <- "2026-09-24"
seasons <- c(2010, 2013, 2017, 2019, 2025, 2026)

pbp_paths <- project_path(
  "data", "raw", "nflverse",
  paste0("pbp_", seasons, "_asof_", snapshot_date, ".rds")
) |>
  set_names(seasons)

schedule_path <- project_path(
  "data", "raw", "nflverse",
  paste0(
    "schedules_", paste(seasons, collapse = "_"),
    "_asof_", snapshot_date, ".rds"
  )
)
```

```{r}
pbp_list <- map(pbp_paths, ~ as_tibble(readRDS(.x)))
df_schedules <- as_tibble(readRDS(schedule_path))
```

## Historical travel context (Table 1)

All five historical London trips followed a road game, and each London game was followed by a bye week. Table 1 summarizes the travel arrangements, arrival timing, opponents, and results.

```{r}
df_historical_context_data <- tibble::tribble(
  ~team_year, ~opponent_location, ~staging_context, ~arrival_nights,
  ~travel_approach, ~pregame_context, ~international_result,
  ~post_rest, ~first_post_result,
  "49ers, 2010", "Denver Broncos; Wembley Stadium, London",
  "Played at Carolina Oct 24; flew directly from Charlotte after the game",
  "Mon Oct 25; time not reported; 6 nights",
  "Early arrival; spent game week in London",
  "49ers 1–6; Broncos 2–5",
  "Won 24–16; +6.0 points",
  "Week 9 bye; 14 days between games",
  "Won 23–20 (OT) vs St Louis; −1.5 points",
  "49ers, 2013", "Jacksonville Jaguars; Wembley Stadium, London",
  "Played at Tennessee Oct 20; direct overnight flight from Nashville",
  "Mon Oct 21, about 13:00; 6 nights",
  "Early arrival; spent game week in London",
  "49ers 5–2; Jaguars 0–7",
  "Won 42–10; +16.0 points",
  "Week 9 bye; 14 days between games",
  "Lost 9–10 vs Carolina; −7.0 points",
  "Rams, 2017", "Arizona Cardinals; Twickenham Stadium, London",
  "Played at Jacksonville Oct 15; stayed there through Thursday",
  "Fri Oct 20, about 09:00; 2 nights",
  "Brief London stay; prepared in eastern US on usual weekly schedule",
  "Rams 4–2; Cardinals 3–3",
  "Won 33–0; +30.0 points",
  "Week 8 bye; 14 days between games",
  "Won 51–17 at New York Giants; +28.5 points",
  "Rams, 2019", "Cincinnati Bengals; Wembley Stadium, London",
  "Played at Atlanta Oct 20; stayed through Thursday practice",
  "Fri Oct 25 morning; exact time not reported; 2 nights",
  "Brief London stay; same eastern-US preparation pattern as 2017",
  "Rams 4–3; Bengals 0–7",
  "Won 24–10; +2.0 points",
  "Week 9 bye; 14 days between games",
  "Lost 12–17 at Pittsburgh; −9.0 points",
  "Rams, 2025", "Jacksonville Jaguars; Wembley Stadium, London",
  "Played at Baltimore Oct 12; stayed and practiced there during game week",
  "<30 hours before kickoff; exact time not reported; 1 night",
  "Brief London stay; prepared in eastern US, aiming to maintain usual routine",
  "Rams 4–2; Jaguars 4–2",
  "Won 35–7; +25.0 points",
  "Week 8 bye; 14 days between games",
  "Won 34–10 vs New Orleans; +10.0 points"
)

```

```{r}
#| label: historical-context-table

(gt_context_table <- historical_context_table(df_historical_context_data))
```

# Football data and outcomes

## Scoring margin and market residual

Scoring margin is the number of points a team scored minus the number it conceded. Market residual compares this margin with the margin expected from the closing point spread. A positive market residual means the team performed better than the spread implied. For example, a team that wins by 20 points after entering as a 3.5-point underdog has a market residual of +23.5 points.

Each game is considered from both teams’ perspectives. Because the point spread in the source data is expressed from the home team’s perspective, its sign is reversed for the away team. Team abbreviations used in older seasons are also matched to the corresponding franchises in the play-by-play data.

```{r}
#| label: team-schedule-definition
#| purl: true

team_schedule <- function(df_schedules) {
  df_res <- df_schedules |>
    filter(game_type == "REG") |>
    mutate(
      across(
        c(home_team, away_team),
        ~ case_when(
          .x == "STL" ~ "LA",
          .x == "OAK" ~ "LV",
          .x == "SD" ~ "LAC",
          .default = .x
        )
      ),
      home_opponent = away_team,
      away_opponent = home_team
    ) |>
    pivot_longer(
      c(home_team, away_team, home_opponent, away_opponent),
      names_to = c("side", ".value"),
      names_sep = "_",
      cols_vary = "slowest"
    ) |>
    transmute(
      season, 
      game_id, 
      team, 
      opponent, 
      week, 
      gameday,
      points_for = if_else(side == "home", home_score, away_score),
      points_against = if_else(side == "home", away_score, home_score),
      expected_margin = if_else(side == "home", spread_line, -spread_line),
      scoring_margin = points_for - points_against,
      performance_vs_market = scoring_margin - expected_margin
    )
  return(df_res)
}
```

## Check the game information

The selected games are checked for the correct teams, opponents, weeks, and final scores, and for the availability of closing point spreads. Scores in the schedule data are also compared with those in the play-by-play data.

The next step confirms that each following game was the team’s first regular-season game after the trip. These games occurred 14 days after each historical London game, 10 days after Melbourne for the 49ers, and 11 days after Melbourne for the Rams.

```{r}
#| label: validate-schedules

df_team_schedules <- team_schedule(df_schedules)

df_schedule_validation <- df_game_registry |>
  left_join(
    df_team_schedules,
    by = join_by(season, team, game_id),
    relationship = "one-to-one"
  )

df_following_games <- df_team_schedules |>
  arrange(season, team, gameday) |>
  mutate(
    next_game_id = lead(game_id),
    next_game_date = lead(as.Date(gameday)),
    .by = c(season, team)
  ) |>
  select(season, team, game_id, next_game_id, next_game_date)

df_episode_validation <- df_schedule_validation |>
  filter(phase == "international") |>
  left_join(df_following_games, by = join_by(season, team, game_id)) |>
  left_join(
    df_game_registry |>
      filter(phase == "return") |>
      select(season, team, registered_return = game_id),
    by = join_by(season, team),
    relationship = "one-to-one"
  ) |>
  mutate(
    days_to_next_game = as.integer(next_game_date - as.Date(gameday)),
    expected_interval = case_when(
      season < 2026 ~ 14,
      team == "SF" ~ 10,
      team == "LA" ~ 11
    )
  )
```

```{r}
#| label: validate-play-by-play

validate_pbp_scores <- function(year) {
  
  df_pbp_year <- pbp_list[[as.character(year)]]
  df_targets <- df_schedule_validation |> filter(season == year)
  
  df_scores <- df_pbp_year |>
    semi_join(df_targets, by = "game_id") |>
    summarise(
      home_team = first(home_team),
      home_score = max(total_home_score),
      away_score = max(total_away_score),
      .by = game_id
    )

  df_res <- df_targets |>
    left_join(df_scores, by = "game_id", relationship = "many-to-one") |>
    mutate(
      pbp_points_for = if_else(team == home_team, home_score, away_score),
      pbp_points_against = if_else(team == home_team, away_score, home_score),
      passed = points_for == pbp_points_for & points_against == pbp_points_against
    ) |>
    select(season, team, game_id, points_for, points_against, passed)
  
  return(df_res)
}

df_pbp_validation <- map(seasons, validate_pbp_scores) |> list_rbind()
df_pbp_validation |> knitr::kable()
```

## Measures of play performance

Expected points added (EPA) describes the change in expected scoring value associated with a play. Offensive EPA per play and defensive EPA allowed per play summarize these values across a game. Success rate is the percentage of plays with positive EPA. Higher offensive values and lower defensive values indicate better performance.

Completion percentage over expected (CPOE) describes completion performance relative to the modeled probability of completing a pass. It is reported in percentage points. Offensive and defensive sack rates describe the percentage of quarterback dropbacks that ended in a sack.

The table below identifies which plays contribute to each season average and whether higher or lower values indicate better performance. Scoring margin and market residual are measured in points, EPA in points per play, and success and sack rates in percentages.

```{r}
#| label: outcome-specification
#| purl: true

df_metric_spec <- tribble(
  ~metric, ~label, ~direction, ~weight, ~standardize,
  "scoring_margin", "Scoring margin", 1, NA, FALSE,
  "performance_vs_market", "Performance versus market", 1, NA, FALSE,
  "off_epa_per_play", "Offensive EPA per play", 1, "offensive_plays", TRUE,
  "def_epa_allowed", "Defensive EPA allowed per play", -1, "defensive_plays", TRUE,
  "off_success_rate", "Offensive success rate", 1, "offensive_plays", TRUE,
  "def_success_rate_allowed", "Defensive success rate allowed", -1, "defensive_plays", TRUE,
  "cpoe", "CPOE", 1, "pass_attempts", FALSE,
  "off_sack_rate", "Sack rate allowed", -1, "offensive_dropbacks", FALSE,
  "def_sack_rate", "Defensive sack rate", 1, "defensive_dropbacks", FALSE
)
```

```{r}
#| label: outcome-guide
df_metric_spec |>
  transmute(
    Outcome = label,
    `Season-average weight` = coalesce(weight, "Equal games"),
    `Better performance` = if_else(direction == 1, "Higher", "Lower"),
    Standardized = if_else(standardize, "Yes", "No")
  ) |>
  knitr::kable()
```

## Which plays are included?

All play-based measures use regular-season plays for which both teams are recorded. Quarterback kneels and spikes are excluded. The remaining requirements depend on the measure:

- **EPA and success rate:** plays identified as runs or passes with an EPA value available. Eligible two-point attempts and overtime plays are included. Plays involving penalties are retained when they meet these requirements, using the EPA recorded after penalty enforcement. A repeated down is treated as a separate play; when a penalty is declined, the original play stands.
- **CPOE:** pass attempts with both completion status and CPOE recorded.
- **Sack rates:** quarterback dropbacks, including scrambles, with sack status recorded.

A team’s defensive results describe what the opposing offense did on those same plays. Completion performance against each defense is also calculated for use in opponent adjustment, although defensive CPOE is not reported as a separate outcome.

```{r}
#| label: game-metrics-definition
#| purl: true

calculate_game_metrics <- function(df_pbp, df_schedules) {
  
  # filter to regular plays
  df_reg_plays <- df_pbp |>
    filter(
      season_type == "REG",
      !coalesce(qb_kneel == 1, FALSE),
      !coalesce(qb_spike == 1, FALSE),
      !is.na(posteam), posteam != "",
      !is.na(defteam), defteam != ""
    )
  
  df_epa_plays <- df_reg_plays |> filter(rush == 1 | pass == 1, !is.na(epa))
  stopifnot(all(df_epa_plays$success == as.numeric(df_epa_plays$epa > 0)))

  df_offense <- df_epa_plays |>
    summarise(
      offensive_plays = n(),
      off_epa_per_play = mean(epa),
      off_success_rate = mean(epa > 0),
      .by = c(season, game_id, posteam, defteam)
    )

  df_passing <- df_reg_plays |>
    filter(pass_attempt == 1, !is.na(cpoe), !is.na(complete_pass)) |>
    summarise(
      pass_attempts = n(),
      cpoe = mean(cpoe),
      .by = c(season, game_id, posteam, defteam)
    )

  df_sacks <- df_reg_plays |>
    filter(qb_dropback == 1, !is.na(sack)) |>
    summarise(
      offensive_dropbacks = n(),
      sacks_allowed = sum(sack),
      off_sack_rate = mean(sack),
      .by = c(season, game_id, posteam, defteam)
    )

  df_offensive_units <- df_offense |>
    left_join(
      df_passing,
      by = join_by(season, game_id, posteam, defteam),
      relationship = "one-to-one"
    ) |>
    left_join(
      df_sacks,
      by = join_by(season, game_id, posteam, defteam),
      relationship = "one-to-one"
    )

  df_defensive_units <- df_offensive_units |>
    transmute(
      season, game_id, team = defteam,
      defensive_plays = offensive_plays,
      def_epa_allowed = off_epa_per_play,
      def_success_rate_allowed = off_success_rate,
      defensive_pass_attempts = pass_attempts,
      def_cpoe_allowed = cpoe,
      defensive_dropbacks = offensive_dropbacks,
      sacks_generated = sacks_allowed,
      def_sack_rate = off_sack_rate
    )

  df_games <- team_schedule(df_schedules) |>
    semi_join(df_offensive_units, by = join_by(season, game_id)) |>
    left_join(
      df_offensive_units |> rename(team = posteam) |> select(-defteam),
      by = join_by(season, game_id, team),
      relationship = "one-to-one"
    ) |>
    left_join(
      df_defensive_units,
      by = join_by(season, game_id, team),
      relationship = "one-to-one"
    )

  return(df_games)
}
```

Game-level performance is calculated for every NFL team in the five comparison seasons. This is necessary because assessing the strength of an opponent requires information about that opponent’s other games. For 2026, only Melbourne and the two following games are needed.

```{r}
#| label: calculate-observed-metrics

reference_seasons <- seasons[seasons != 2026]
df_league_game_metrics <- map_dfr(reference_seasons, function(year) {
  df_pbp_year <- pbp_list[[as.character(year)]]
  calculate_game_metrics(df_pbp_year, df_schedules)
}) 

df_historical_registry <- df_game_registry |>
  filter(season < 2026) |>
  select(season, team, game_id, phase)

df_melbourne_registry <- df_game_registry |>
  filter(season == 2026) |>
  select(season, team, game_id, phase)

df_melbourne_observed <- pbp_list[["2026"]] |>
  semi_join(df_melbourne_registry, by = "game_id") |>
  calculate_game_metrics(df_schedules)
```

# Opponent-adjusted analyses

Teams face opponents of differing strength. The opponent-adjusted analyses describe how a game’s performance compares after allowing for the opponent’s usual performance relative to the league:

$$\text{Adjusted value} = \text{Observed value}
- (\text{Opponent average} - \text{League average}).$$

An offensive measure is compared with the opponent’s usual defensive performance; a defensive measure is compared with the opponent’s usual offensive performance. For example, offensive EPA is adjusted downward after facing a defense that usually allows more EPA than the league average, and upward after facing a defense that usually allows less.

This adjustment is applied to the seven play-based measures. Scoring margin and market residual are not opponent-adjusted.

```{r}
#| label: opponent-specification
#| purl: true

df_opponent_spec <- tribble(
  ~metric, ~opponent_metric, ~opponent_weight, ~league_metric,
  "off_epa_per_play", "def_epa_allowed", "defensive_plays", "off_epa_per_play",
  "def_epa_allowed", "off_epa_per_play", "offensive_plays", "off_epa_per_play",
  "off_success_rate", "def_success_rate_allowed", "defensive_plays", "off_success_rate",
  "def_success_rate_allowed", "off_success_rate", "offensive_plays", "off_success_rate",
  "cpoe", "def_cpoe_allowed", "defensive_pass_attempts", "cpoe",
  "off_sack_rate", "def_sack_rate", "defensive_dropbacks", "off_sack_rate",
  "def_sack_rate", "off_sack_rate", "offensive_dropbacks", "off_sack_rate"
)

df_opponent_spec <- df_opponent_spec |> 
  left_join(
    df_metric_spec |> select(league_metric = metric, league_weight = weight),
    by = "league_metric",
    relationship = "many-to-one"
  )
```

For the historical games and the games in the 2025 comparison season, the opponent’s average is calculated from its other regular-season games that year. The game being assessed is excluded from that opponent average. The league average includes all regular-season games in the same year.

Melbourne and its following games occurred near the start of the 2026 season. Their adjustments therefore use the complete 2025 season for both opponent and league averages.

These averages account for the number of opportunities contributing to each measure: eligible plays for EPA and success rate, pass attempts for CPOE, and quarterback dropbacks for sack rates.

```{r}
#| label: opponent-adjustment-definitions
#| purl: true

adjust_one_metric <- function(metric, opponent_metric, opponent_weight, league_metric,
                              league_weight, df_games, df_league_games, reference_season) {
  
  # rename for clarity
  df_reference <- df_league_games
  if (!is.null(reference_season)) {
    df_reference <- df_reference |> filter(season == reference_season)
  }

  df_league_ratings <- df_reference |>
    summarise(
      league_value = weighted.mean(.data[[league_metric]], .data[[league_weight]]),
      .by = season
    )

  df_opponent_games <- df_reference |>
    transmute(
      season, 
      game_id, 
      opponent = team,
      numerator = .data[[opponent_metric]] * .data[[opponent_weight]],
      denominator = .data[[opponent_weight]]
    )

  if (is.null(reference_season)) {
    df_opponent_ratings <- df_opponent_games |>
      mutate(
        opponent_value = (sum(numerator) - numerator) / (sum(denominator) - denominator),
        opponent_reference_n_games = n() - 1,
        .by = c(season, opponent)
      ) |>
      select(season, game_id, opponent, opponent_value, opponent_reference_n_games)

    df_matched_ratings <- df_games |>
      select(season, team, game_id, opponent, observed_value = all_of(metric)) |>
      left_join(
        df_opponent_ratings,
        by = join_by(season, game_id, opponent),
        relationship = "many-to-one"
      ) |>
      left_join(df_league_ratings, by = "season", relationship = "many-to-one")
    
  } else {
    df_opponent_ratings <- df_opponent_games |>
      summarise(
        opponent_value = sum(numerator) / sum(denominator),
        opponent_reference_n_games = n(),
        .by = opponent
      )

    df_matched_ratings <- df_games |>
      select(season, team, game_id, opponent, observed_value = all_of(metric)) |>
      left_join(df_opponent_ratings, by = "opponent", relationship = "many-to-one") |>
      mutate(league_value = df_league_ratings$league_value)
  }
  
   df_res <- df_matched_ratings |>
      transmute(
        season, 
        team, 
        game_id, 
        opponent_reference_n_games,
        metric = paste0("adj_", metric),
        adjusted_value = observed_value - (opponent_value - league_value)
    )
 
  return(df_res)
}

adjust_for_opponent <- function(df_games, df_league_games, reference_season = NULL) {
  
  df_adjusted_metrics <- pmap_dfr(
    df_opponent_spec,
    adjust_one_metric,
    df_games = df_games,
    df_league_games = df_league_games,
    reference_season = reference_season
  ) |> 
    pivot_wider(names_from = metric, values_from = adjusted_value)

  df_res <- df_games |>
    left_join(
      df_adjusted_metrics,
      by = join_by(season, team, game_id),
      relationship = "one-to-one"
    ) |>
    relocate(opponent_reference_n_games, .after = adj_off_epa_per_play)
  
  return(df_res)
}
```

```{r}
#| label: calculate-adjusted-metrics

df_team_game_metrics <- df_league_game_metrics |>
  semi_join(df_reference_team_seasons, by = join_by(season, team)) |>
  adjust_for_opponent(df_league_game_metrics) |>
  arrange(season, team, gameday)

df_historical_game_metrics <- df_historical_registry |>
  left_join(
    df_team_game_metrics,
    by = join_by(season, team, game_id),
    relationship = "one-to-one"
  )

df_melbourne_game_metrics <- df_melbourne_registry |>
  left_join(
    df_melbourne_observed,
    by = join_by(season, team, game_id),
    relationship = "one-to-one"
  ) |>
  adjust_for_opponent(df_league_game_metrics, reference_season = 2025)
```

# Season references and standardized comparisons

Each selected game is compared with a full season for the same team. Historical games use the corresponding season, while Melbourne and its following games use 2025. Unadjusted game values are compared with unadjusted season values, and opponent-adjusted game values with opponent-adjusted season values.

The season comparison differs according to the question being asked:

| Comparison | How games contribute | Games included |
|------------------------|------------------------|------------------------|
| Season averages reported in the original units | Games contribute equally to scoring margin and market residual. Play-based measures account for the number of eligible plays, pass attempts, or dropbacks. | The full comparison season, including the game being assessed in historical analyses. |
| Standardized differences and counts of better or worse games | Each game contributes equally. | Historical comparisons exclude the game being assessed. Melbourne and its following games use all 17 games from 2025. |
| Historical scoring and market comparisons in Figure 3 | Each game contributes equally. | The season excluding both the London game and its first following game. |

Standardized differences place offensive and defensive EPA and success rates on a common scale:

$$\text{Standardized difference} = \text{Direction}\times
\frac{\text{Game value}-\text{Comparison mean}}
{\text{Comparison standard deviation}}.$$

The mean and sample standard deviation are calculated from the game-level values. The standard deviation therefore describes how much performance varied between games.

The direction is set so that positive values always indicate better performance: +1 for measures where higher values are better, and −1 for measures where lower values are better. A standardized difference of +1 means that performance was one standard deviation more favorable than the comparison mean.

CPOE and sack rates are reported in their original units. They describe fewer opportunities than EPA and success rate, and a small number of sacks can substantially change a game’s sack rate.

## Calculate the season comparisons

The calculations below bring together each game’s performance and the appropriate season comparison. They produce both the season averages in the original units and the standardized differences for the four EPA and success-rate measures.

```{r}
df_comparison_spec <- bind_rows(
  df_metric_spec |> mutate(column = metric, adjustment = "Unadjusted"),
  df_metric_spec |>
    semi_join(df_opponent_spec, by = "metric") |>
    mutate(column = paste0("adj_", metric), adjustment = "Opponent adjusted")
) |> 
  arrange(match(metric, df_metric_spec$metric), adjustment == "Opponent adjusted")
```

```{r}
#| label: comparison-table-definitions
#| purl: true

game_metrics_long <- function(df_games) {
  
  weight_columns <- df_metric_spec$weight |> na.omit() |> unique() |> as.character()

  df_denominators <- df_games |>
    select(season, team, game_id, all_of(weight_columns)) |>
    pivot_longer(
      all_of(weight_columns),
      names_to = "weight",
      values_to = "opportunities"
    )

  df_values <- df_games |>
    select(season, team, game_id, any_of("phase"), all_of(df_comparison_spec$column)) |>
    pivot_longer(
      all_of(df_comparison_spec$column),
      names_to = "column",
      values_to = "value"
    ) |>
    left_join(df_comparison_spec, by = "column", relationship = "many-to-one") |>
    left_join(
      df_denominators,
      by = join_by(season, team, game_id, weight),
      relationship = "many-to-one"
    ) |>
    mutate(opportunities = if_else(is.na(weight), 1, opportunities))
  
  return(df_values)
}

compare_with_reference <- function(df_focal_games, df_reference_games) {
  
  df_reference_values <- game_metrics_long(df_reference_games)

  # Full-season averages in native units retain all reference-season games.
  df_season_means <- df_reference_values |>
    summarise(
      season_mean = weighted.mean(value, opportunities),
      .by = c(season, team, metric, adjustment)
    )

  df_focal_values <- game_metrics_long(df_focal_games) |>
    mutate(reference_season = if_else(season == 2026L, 2025L, season)) |>
    select(
      season, team, game_id, phase, metric, label, direction, adjustment,
      standardize, reference_season, target_value = value
    )

  # This deliberate many-to-many join pairs each focal outcome with every
  # matching reference game, before excluding the focal game itself.
  df_reference_pairs <- df_focal_values |>
    left_join(
      df_reference_values |>
        select(
          reference_season = season, team, metric, adjustment,
          reference_game_id = game_id, reference_value = value
        ),
      by = join_by(reference_season, team, metric, adjustment),
      relationship = "many-to-many"
    ) |>
    filter(reference_game_id != game_id)

  df_distributions <- df_reference_pairs |>
    summarise(
      reference_n_games = n(),
      reference_mean = mean(reference_value),
      reference_sd = sd(reference_value),
      games_exceeded = sum(direction * reference_value < direction * target_value),
      games_tied = sum(reference_value == target_value),
      .by = c(season, team, game_id, metric, adjustment)
    )

  df_comparisons <- df_focal_values |>
    left_join(
      df_season_means,
      by = join_by(reference_season == season, team, metric, adjustment),
      relationship = "many-to-one"
    ) |>
    left_join(
      df_distributions,
      by = join_by(season, team, game_id, metric, adjustment),
      relationship = "one-to-one"
    ) |>
    mutate(
      standardized_score = if_else(
        standardize,
        direction * (target_value - reference_mean) / reference_sd,
        NA_real_
      )
    ) |>
    select(
      season, team, game_id, phase, metric, label, direction, adjustment,
      target_value, season_mean, reference_n_games, reference_mean, reference_sd,
      games_exceeded, games_tied, standardized_score
    )

  return(df_comparisons)
}
```

## Compare performance with individual season games

The analysis also counts how many games in the comparison season had less favorable performance than the game being assessed. Equal values are counted separately. These counts support descriptions such as “better than 16 of 17 games” and show where a performance falls among that team’s season results.

The season averages used in Figure 3 are calculated separately because they exclude both the London game and its first following game.

```{r}
#| label: calculate-comparisons

df_melbourne_comparisons <- compare_with_reference(
  df_melbourne_game_metrics, 
  df_team_game_metrics
)

df_historical_comparisons <- compare_with_reference(
  df_historical_game_metrics, 
  df_team_game_metrics
)

df_historical_score_context <- df_team_game_metrics |>
  semi_join(df_historical_registry, by = join_by(season, team)) |>
  anti_join(df_historical_registry, by = join_by(season, team, game_id)) |>
  summarise(
    across(c(scoring_margin, performance_vs_market), mean),
    n_games = n(),
    .by = c(season, team)
  )

df_team_game_metrics |>
  summarise(reference_games = n(), .by = c(season, team)) |>
  knitr::kable()
```

# Results

Results are presented for Melbourne and the first following games, then for the historical London episodes. The tables show the game values, season averages, and comparisons with individual season games. Success and sack rates are displayed as percentages, and CPOE as percentage points.

The “season mean” column gives the average in the measure’s original units. Standardized differences use the separate, equally weighted game comparisons described above.

```{r}
#| label: results-display

format_outcome <- function(metric, value) {
  formatted_values <- case_when(
    str_detect(metric, "success|sack") ~ sprintf("%.1f%%", 100 * value),
    metric == "cpoe" ~ sprintf("%+.1f pp", value),
    str_detect(metric, "epa") ~ sprintf("%+.3f", value),
    .default = sprintf("%+.1f", value)
  )
  return(formatted_values)
}

show_comparisons <- function(df_comparisons) {
  df_display <- df_comparisons |>
    transmute(
      season, team, phase, outcome = label, adjustment,
      observed = format_outcome(metric, target_value),
      season_mean = format_outcome(metric, season_mean),
      reference_n_games, games_exceeded, games_tied,
      standardized_difference = round(standardized_score, 2)
    )
  comparison_table <- rmarkdown::paged_table(df_display)
  return(comparison_table)
}

offensive_outcomes <- c("off_epa_per_play", "off_success_rate", "cpoe", "off_sack_rate")
defensive_outcomes <- c("def_epa_allowed", "def_success_rate_allowed", "def_sack_rate")
df_melbourne_matchup <- df_melbourne_comparisons |> filter(phase == "international")
```

## Melbourne episode

### Game summary

The first table shows the final score, expected margin, and market residual from each team’s perspective. The next table compares the scoring margin and market residual with the same team’s 2025 season.

```{r}
#| label: melbourne-game-summary

df_melbourne_game_metrics |>
  filter(phase == "international") |>
  select(
    team,
    opponent,
    points_for,
    points_against,
    expected_margin,
    scoring_margin,
    performance_vs_market
  ) |> 
  knitr::kable()

df_melbourne_matchup |>
  filter(metric %in% c("scoring_margin", "performance_vs_market")) |>
  show_comparisons()
```

### 49ers offense versus Rams defense

San Francisco’s offensive EPA and success rate also describe the EPA and success rate allowed by the Rams’ defense. Although these are the same plays, each team is compared with its own 2025 performance. The tables show these comparisons with and without opponent adjustment, together with CPOE and sack rates.

```{r}
#| label: san-francisco-offense

df_melbourne_matchup |>
  filter(
    (team == "SF" & metric %in% offensive_outcomes) |
      (team == "LA" & metric %in% defensive_outcomes)
  ) |>
  show_comparisons()
```

### Rams offense versus 49ers defense

The same comparisons are then made for the Rams’ offense and the 49ers’ defense. The play, pass-attempt, and dropback counts show how many opportunities contributed to each team’s measures.

```{r}
#| label: los-angeles-offense

df_melbourne_matchup |>
  filter(
    (team == "LA" & metric %in% offensive_outcomes) |
      (team == "SF" & metric %in% defensive_outcomes)
  ) |>
  show_comparisons()

df_melbourne_game_metrics |>
  filter(phase == "international") |>
  select(
    team,
    offensive_plays,
    pass_attempts,
    offensive_dropbacks,
    sacks_allowed,
    defensive_plays,
    defensive_dropbacks,
    sacks_generated
  ) |> 
  knitr::kable()
```

### Melbourne performance (Figure 2)

Figure 2 brings together the game outcome relative to the closing spread, EPA and success rates in their original units, and their standardized differences from 2025 performance. Positive standardized values indicate better performance. CPOE and sack rates are reported in the preceding tables.

```{r}
#| label: melbourne-performance
#| fig-width: 9
#| fig-height: 9.4

melbourne_figure <- build_melbourne_performance_figure(
  df_melbourne_matchup |> filter(adjustment == "Unadjusted"),
  df_melbourne_matchup |> filter(adjustment == "Opponent adjusted"),
  df_melbourne_game_metrics |> filter(phase == "international")
)
save_paper_figure(melbourne_figure, "melbourne_outcome_performance", 9, 9.4)
melbourne_figure
```

### First games after Melbourne

Each team’s first game after Melbourne is compared with the same 2025 season used for the Melbourne analysis. The opponent adjustments also continue to use 2025. This allows performance in Melbourne and in the next game to be read against the same prior-season comparisons.

```{r}
#| label: post-melbourne-results

df_melbourne_game_metrics |>
  filter(phase == "return") |>
  select(team, opponent, gameday, points_for, points_against, performance_vs_market) |>
  knitr::kable()

df_melbourne_comparisons |>
  filter(phase == "return") |>
  show_comparisons()
```

## Historical episodes

### Early-arriving 49ers

The 49ers’ 2010 and 2013 London games and their first following games are compared with San Francisco’s performance in the corresponding season. The tables show scoring and market outcomes alongside play-based performance, with and without opponent adjustment where applicable.

```{r}
#| label: historical-san-francisco

df_historical_comparisons |>
  filter(team == "SF") |>
  show_comparisons()
```

### Short-stay Rams

The Rams’ 2017, 2019, and 2025 episodes are examined in the same way. Each London game and its first following game are compared with the Rams’ performance in that year.

```{r}
#| label: historical-los-angeles

df_historical_comparisons |>
  filter(team == "LA") |>
  show_comparisons()
```

### Historical scoring and market context (Figure 3)

Figure 3 compares scoring margin and market residual in the London and following games with the rest of the season. Both games are excluded from these season averages. The table below shows the averages and the number of remaining games used to calculate them.

```{r}
#| label: historical-score-context

df_historical_score_context |> knitr::kable(digits = 2)
```

```{r}
#| label: historical-outcomes
#| fig-width: 9
#| fig-height: 5.85

historical_figures_list <- build_historical_figures(
  df_historical_game_metrics,
  df_historical_score_context,
  df_historical_comparisons
)

save_paper_figure(
  historical_figures_list$outcomes, 
  "historical_score_market_baselines", 
  9, 5.85)

save_paper_figure(
  historical_figures_list$performance,
  "historical_international_return_play_metrics",
  9, 6.25
)
historical_figures_list$outcomes
```

### Historical standardized performance (Figure 4)

Figure 4 compares offensive and defensive EPA and success rates across the historical episodes. Each game is compared with the other games in the same team’s season. The bars show results with and without opponent adjustment; positive values indicate better performance than the corresponding season comparison.

```{r}
#| label: historical-performance
#| fig-width: 9
#| fig-height: 6.25

historical_figures_list$performance
```

# Save and check the results

## Save the analysis results

The game results, season comparisons, and checks of game scores are saved for use with the manuscript.

```{r}
#| label: export-analysis
#| code-fold: true

inputs_list <- list(
  snapshot_date = snapshot_date,
  pbp_paths = pbp_paths,
  schedules = df_schedules,
  game_registry = df_game_registry,
  reference_team_seasons = df_reference_team_seasons
)

historical_results_list <- list(
  snapshot_date = snapshot_date,
  league_game_metrics = df_league_game_metrics,
  team_game_metrics = df_team_game_metrics,
  game_metrics = df_historical_game_metrics,
  comparisons = df_historical_comparisons,
  context = df_historical_score_context
)

melbourne_results_list <- list(
  snapshot_date = snapshot_date,
  game_metrics = df_melbourne_game_metrics,
  comparisons = df_melbourne_comparisons
)

analysis_exports_list <- list(
  game_validation = df_pbp_validation,
  historical_game_metrics = df_historical_game_metrics,
  historical_reference_comparisons = df_historical_comparisons,
  historical_score_market_context = df_historical_score_context,
  melbourne_game_metrics = df_melbourne_game_metrics,
  melbourne_reference_comparisons = df_melbourne_comparisons
)

dir.create(project_path("data", "derived"), recursive = TRUE, showWarnings = FALSE)
saveRDS(inputs_list, project_path("data", "derived", "validated_inputs.rds"))
saveRDS(historical_results_list, project_path("data", "derived", "historical_results.rds"))
saveRDS(melbourne_results_list, project_path("data", "derived", "melbourne_results.rds"))

iwalk(analysis_exports_list, function(df_data, name) {
  path <- project_path("data", "derived", paste0(name, ".csv"))
  write_csv(df_data, path, na = "")
})
```

## Save the figures and Table 1

The figures shown above are saved as PNG images for the manuscript. Table 1 is saved as an image and a Word document. The Melbourne travel figure is maintained separately.

```{r}
#| label: export-context-table
#| code-fold: true

dir.create(project_path("tables"), recursive = TRUE, showWarnings = FALSE)
gt::gtsave(
  gt_context_table,
  project_path("tables", "table_1_historical_international_context.png"),
  vwidth = 1300, vheight = 900, expand = 10
)
save_historical_context_docx(
  gt_context_table,
  project_path("tables", "table_1_historical_international_context.docx")
)
```
