---
title: "NFL Melbourne 2026: reproducible analysis"
format:
  html:
    code-fold: false
    toc: true
editor_options:
  chunk_output_type: console
---

# Setup

This notebook calculates game-level performance, opponent adjustments, and season comparisons for the selected Melbourne and London episodes. Each episode includes the international game and the team’s first following regular-season game.

The analysis contains 14 team-game observations: ten from the historical London episodes and four from Melbourne and its following games. Melbourne contributes an observation for each team.

## Packages and paths

```{r}
#| label: setup

suppressPackageStartupMessages({
  library(tidyverse)
  library(patchwork)
})
```

```{r}
# Resolve paths from the project root or the notebooks directory.
project_path <- function(...) {
  root <- if (file.exists("_quarto.yml")) "." else ".."
  
  if (!file.exists(file.path(root, "_quarto.yml"))) {
    stop("Open nfl-melbourne-2026-travel-performance.Rproj before running the analysis.")
  }
  
  return(file.path(normalizePath(root), ...))
}
```

## Study games

The table below defines the selected international team-games. Melbourne appears twice because both teams are included. Game details and each team’s first following game are derived from the schedule.

```{r}
df_study_games <- tribble(
  ~team, ~game_id,
  "SF", "2010_08_DEN_SF",
  "SF", "2013_08_SF_JAX",
  "LA", "2017_07_ARI_LA",
  "LA", "2019_08_CIN_LA",
  "LA", "2025_07_LA_JAX",
  "SF", "2026_01_SF_LA",
  "LA", "2026_01_SF_LA"
)
```

## Load data

Set `data_source` to `"archive"` to reproduce the submitted analysis using the September 24, 2026 snapshots, or to `"nflverse"` to download current data. Both modes load each season once; fresh downloads may contain revisions.

Historical comparisons use the corresponding season. Melbourne and its following games use `melbourne_reference_season`, set to 2025.

```{r}
#| label: load-data

data_source <- "archive"  # "archive" or "nflverse"
seasons <- c(2010, 2013, 2017, 2019, 2025, 2026)
melbourne_reference_season <- 2025

if (data_source == "archive") {
  snapshot_date <- "2026-09-24"
  retrieved_at <- NA
  output_root <- project_path()

  pbp_paths <- project_path(
    "data", "raw", "nflverse",
    paste0("pbp_", seasons, "_asof_", snapshot_date, ".rds")
  ) |> set_names(seasons)

  schedule_path <- project_path(
    "data", "raw", "nflverse",
    paste0("schedules_", paste(seasons, collapse = "_"),  "_asof_", snapshot_date, ".rds")
  )

  pbp_list <- map(pbp_paths, ~ as_tibble(readRDS(.x)))
  df_schedules <- as_tibble(readRDS(schedule_path))
  
} else if (data_source == "nflverse") {
  # Clear cached responses so this run requests the current nflverse files.
  nflreadr::clear_cache()
  pbp_list <- map(set_names(seasons), ~ as_tibble(nflreadr::load_pbp(.x)))
  df_schedules <- as_tibble(nflreadr::load_schedules(seasons))

  snapshot_date <- NA
  pbp_paths <- NULL
  schedule_path <- NULL
  download_time <- Sys.time()
  retrieved_at <- format(download_time, "%Y-%m-%d %H:%M:%S UTC", tz = "UTC")
  output_root <- project_path(
    "outputs", 
    "nflverse",
    format(download_time, "%Y%m%dT%H%M%SZ", tz = "UTC")
    
  )
} else {
  stop('Choose data_source = "archive" or "nflverse".')
}

derived_directory <- file.path(output_root, "data", "derived")
figure_directory <- file.path(output_root, "figures")
table_directory <- file.path(output_root, "tables")

df_data_source <- tibble(
  source = data_source,
  seasons = paste(seasons, collapse = ", "),
  archive_date = snapshot_date,
  retrieved_at = retrieved_at,
  output_directory = output_root
)
```

## Helper functions

The following chunks load the presentation helpers and define the schedule, game-metric, opponent-adjustment, and comparison functions used below.

```{r}
#| code-fold: true

# Load the figure and table presentation helpers.
source(project_path("R", "figure_helpers.R"))
source(project_path("R", "historical_figures.R"))
source(project_path("R", "melbourne_performance_figure.R"))
source(project_path("R", "contextual_tables.R"))
```

```{r}
#| label: team-schedule-definition
#| code-fold: true

team_schedule <- function(df_schedules) {
  df_res <- df_schedules |>
    filter(game_type == "REG") |>
    mutate(
      across(
        c(home_team, away_team),
        ~ case_when(
          .x == "STL" ~ "LA",
          .x == "OAK" ~ "LV",
          .x == "SD" ~ "LAC",
          .default = .x
        )
      ),
      home_opponent = away_team,
      away_opponent = home_team
    ) |>
    pivot_longer(
      c(home_team, away_team, home_opponent, away_opponent),
      names_to = c("side", ".value"),
      names_sep = "_",
      cols_vary = "slowest"
    ) |>
    transmute(
      season, 
      game_id, 
      team, 
      opponent, 
      week, 
      gameday,
      points_for = if_else(side == "home", home_score, away_score),
      points_against = if_else(side == "home", away_score, home_score),
      expected_margin = if_else(side == "home", spread_line, -spread_line),
      scoring_margin = points_for - points_against,
      performance_vs_market = scoring_margin - expected_margin
    )
  return(df_res)
}
```

```{r}
#| label: game-metrics-definition
#| code-fold: true

calculate_game_metrics <- function(df_reg_plays, df_team_schedules) {
  
  df_epa_plays <- df_reg_plays |> filter(rush == 1 | pass == 1, !is.na(epa))
  
  df_offense <- df_epa_plays |>
    summarise(
      offensive_plays = n(),
      off_epa_per_play = mean(epa),
      off_success_rate = mean(epa > 0),
      .by = c(season, game_id, posteam, defteam)
    )
  
  df_passing <- df_reg_plays |>
    filter(pass_attempt == 1, !is.na(cpoe), !is.na(complete_pass)) |>
    summarise(
      pass_attempts = n(),
      cpoe = mean(cpoe),
      .by = c(season, game_id, posteam, defteam)
    )
  
  df_sacks <- df_reg_plays |>
    filter(qb_dropback == 1, !is.na(sack)) |>
    summarise(
      offensive_dropbacks = n(),
      sacks_allowed = sum(sack),
      off_sack_rate = mean(sack),
      .by = c(season, game_id, posteam, defteam)
    )
  
  df_offensive_units <- df_offense |>
    left_join(
      df_passing,
      by = join_by(season, game_id, posteam, defteam),
      relationship = "one-to-one"
    ) |>
    left_join(
      df_sacks,
      by = join_by(season, game_id, posteam, defteam),
      relationship = "one-to-one"
    )
  
  df_defensive_units <- df_offensive_units |>
    transmute(
      season, game_id, team = defteam,
      defensive_plays = offensive_plays,
      def_epa_allowed = off_epa_per_play,
      def_success_rate_allowed = off_success_rate,
      defensive_pass_attempts = pass_attempts,
      def_cpoe_allowed = cpoe,
      defensive_dropbacks = offensive_dropbacks,
      sacks_generated = sacks_allowed,
      def_sack_rate = off_sack_rate
    )
  
  df_games <- df_team_schedules |>
    semi_join(df_offensive_units, by = join_by(season, game_id)) |>
    left_join(
      df_offensive_units |> rename(team = posteam) |> select(-defteam),
      by = join_by(season, game_id, team),
      relationship = "one-to-one"
    ) |>
    left_join(
      df_defensive_units,
      by = join_by(season, game_id, team),
      relationship = "one-to-one"
    )
  
  return(df_games)
}
```

```{r}
#| label: opponent-adjustment-definitions
#| code-fold: true

adjust_one_metric <- function(metric, opponent_metric, opponent_weight, league_metric,
                              league_weight, df_games, df_league_games, reference_season) {
  
  # Use the specified reference season when supplied; otherwise retain all seasons.
  df_reference <- df_league_games
  if (!is.null(reference_season)) {
    df_reference <- df_reference |> filter(season == reference_season)
  }
  
  df_league_ratings <- df_reference |>
    summarise(
      league_value = weighted.mean(.data[[league_metric]], .data[[league_weight]]),
      .by = season
    )
  
  df_opponent_games <- df_reference |>
    transmute(
      season, 
      game_id, 
      opponent = team,
      numerator = .data[[opponent_metric]] * .data[[opponent_weight]],
      denominator = .data[[opponent_weight]]
    )
  
  if (is.null(reference_season)) {
    df_opponent_ratings <- df_opponent_games |>
      mutate(
        opponent_value = (sum(numerator) - numerator) / (sum(denominator) - denominator),
        opponent_reference_n_games = n() - 1,
        .by = c(season, opponent)
      ) |>
      select(season, game_id, opponent, opponent_value, opponent_reference_n_games)
    
    df_matched_ratings <- df_games |>
      select(season, team, game_id, opponent, observed_value = all_of(metric)) |>
      left_join(
        df_opponent_ratings,
        by = join_by(season, game_id, opponent),
        relationship = "many-to-one"
      ) |>
      left_join(df_league_ratings, by = "season", relationship = "many-to-one")
    
  } else {
    df_opponent_ratings <- df_opponent_games |>
      summarise(
        opponent_value = sum(numerator) / sum(denominator),
        opponent_reference_n_games = n(),
        .by = opponent
      )
    
    df_matched_ratings <- df_games |>
      select(season, team, game_id, opponent, observed_value = all_of(metric)) |>
      left_join(df_opponent_ratings, by = "opponent", relationship = "many-to-one") |>
      mutate(league_value = df_league_ratings$league_value)
  }
  
  df_res <- df_matched_ratings |>
    transmute(
      season, 
      team, 
      game_id, 
      opponent_reference_n_games,
      metric = paste0("adj_", metric),
      adjusted_value = observed_value - (opponent_value - league_value)
    )
  
  return(df_res)
}

adjust_for_opponent <- function(df_games, df_league_games, reference_season = NULL) {
  
  df_adjusted_metrics <- pmap_dfr(
    df_opponent_spec,
    adjust_one_metric,
    df_games = df_games,
    df_league_games = df_league_games,
    reference_season = reference_season
  ) |> 
    pivot_wider(names_from = metric, values_from = adjusted_value)
  
  df_res <- df_games |>
    left_join(
      df_adjusted_metrics,
      by = join_by(season, team, game_id),
      relationship = "one-to-one"
    ) |>
    relocate(opponent_reference_n_games, .after = adj_off_epa_per_play)
  
  return(df_res)
}
```

```{r}
#| label: comparison-table-definitions
#| code-fold: true

game_metrics_long <- function(df_games) {
  
  weight_columns <- df_metric_spec$weight |> na.omit() |> unique() |> as.character()
  
  df_denominators <- df_games |>
    select(season, team, game_id, all_of(weight_columns)) |>
    pivot_longer(
      all_of(weight_columns),
      names_to = "weight",
      values_to = "opportunities"
    )
  
  df_values <- df_games |>
    select(season, team, game_id, any_of("phase"), all_of(df_comparison_spec$column)) |>
    pivot_longer(
      all_of(df_comparison_spec$column),
      names_to = "column",
      values_to = "value"
    ) |>
    left_join(df_comparison_spec, by = "column", relationship = "many-to-one") |>
    left_join(
      df_denominators,
      by = join_by(season, team, game_id, weight),
      relationship = "many-to-one"
    ) |>
    mutate(opportunities = if_else(is.na(weight), 1, opportunities))
  
  return(df_values)
}

compare_with_reference <- function(df_focal_games, df_reference_games) {
  
  df_reference_values <- game_metrics_long(df_reference_games)
  
  # Full-season averages in native units retain all reference-season games.
  df_season_means <- df_reference_values |>
    summarise(
      season_mean = weighted.mean(value, opportunities),
      .by = c(season, team, metric, adjustment)
    )
  
  df_focal_values <- game_metrics_long(df_focal_games) |>
    mutate(reference_season = if_else(season == 2026, melbourne_reference_season, season)) |>
    select(
      season, team, game_id, phase, metric, label, direction, adjustment,
      standardize, reference_season, target_value = value
    )
  
  # This deliberate many-to-many join pairs each focal outcome with every
  # matching reference game, before excluding the focal game itself.
  df_reference_pairs <- df_focal_values |>
    left_join(
      df_reference_values |>
        select(
          reference_season = season, team, metric, adjustment,
          reference_game_id = game_id, reference_value = value
        ),
      by = join_by(reference_season, team, metric, adjustment),
      relationship = "many-to-many"
    ) |>
    filter(reference_game_id != game_id)
  
  df_distributions <- df_reference_pairs |>
    summarise(
      reference_n_games = n(),
      reference_mean = mean(reference_value),
      reference_sd = sd(reference_value),
      games_exceeded = sum(direction * reference_value < direction * target_value),
      games_tied = sum(reference_value == target_value),
      .by = c(season, team, game_id, metric, adjustment)
    )
  
  df_comparisons <- df_focal_values |>
    left_join(
      df_season_means,
      by = join_by(reference_season == season, team, metric, adjustment),
      relationship = "many-to-one"
    ) |>
    left_join(
      df_distributions,
      by = join_by(season, team, game_id, metric, adjustment),
      relationship = "one-to-one"
    ) |>
    mutate(
      standardized_score = if_else(
        standardize,
        direction * (target_value - reference_mean) / reference_sd,
        NA
      )
    ) |>
    select(
      season, team, game_id, phase, metric, label, direction, adjustment,
      target_value, season_mean, reference_n_games, reference_mean, reference_sd,
      games_exceeded, games_tied, standardized_score
    )
  
  return(df_comparisons)
}
```

# Game selection and performance

## Study and reference games

Each schedule record is converted into two team-level records. Scores and closing expected margins are expressed from each team’s perspective, and historical team abbreviations are matched across datasets.

The schedule is ordered by date within each team and season to identify the first game after each international game, including when a bye week intervenes. The first table shows these games and the elapsed days between them.

```{r}
#| label: select-study-games

df_team_schedules <- team_schedule(df_schedules)

df_following_games <- df_team_schedules |>
  arrange(season, team, gameday) |>
  mutate(
    next_game_id = lead(game_id),
    next_game_date = lead(as.Date(gameday)),
    .by = c(season, team)
  )

df_study_schedule <- df_study_games |>
  left_join(df_following_games, by = join_by(team, game_id), relationship = "one-to-one")
```

```{r}
#| code-fold: true

df_study_schedule |>
  transmute(
    season, team, game_id, next_game_id, gameday, next_game_date,
    days_to_next_game = as.integer(next_game_date - as.Date(gameday))
  ) |>
  knitr::kable()
```

The registry combines the international and following games. The second table shows their opponents, weeks, and closing expected margins. Reference team-seasons are then derived from the selected episodes.

```{r}
df_game_registry <- bind_rows(
  df_study_schedule |> transmute(season, team, phase = "international", game_id),
  df_study_schedule |> transmute(season, team, phase = "return", game_id = next_game_id)
) |>
  arrange(season, phase == "return") |>
  left_join(df_team_schedules, by = join_by(season, team, game_id), relationship = "one-to-one")
```

```{r}
#| code-fold: true

df_game_registry |>
  select(season, team, phase, game_id, opponent, week, expected_margin) |>
  knitr::kable()
```

```{r}
df_reference_team_seasons <- df_study_schedule |>
  transmute(season = if_else(season == 2026, melbourne_reference_season, season), team) |>
  distinct()
```

## Performance measures

**Scoring margin** is points scored minus points conceded. **Market residual** is scoring margin minus the margin implied by the closing point spread; positive values indicate performance above that expectation.

**Expected points added (EPA)** describes the change in expected scoring value associated with a play. **Success rate** is the proportion of eligible plays with positive EPA. Higher offensive EPA and success rates, and lower defensive values, indicate better performance.

**Completion percentage over expected (CPOE)** measures completion performance relative to modeled completion probabilities. **Sack rates** are the proportions of quarterback dropbacks ending in sacks; lower offensive and higher defensive rates indicate better performance.

The table identifies the weights used for season averages, the favorable direction of each outcome, and which outcomes are standardized. Success and sack rates are stored as proportions and displayed as percentages; CPOE is reported in percentage points.

```{r}
#| label: outcome-specification
#| code-fold: true

df_metric_spec <- tribble(
  ~metric, ~label, ~direction, ~weight, ~standardize,
  "scoring_margin", "Scoring margin", 1, NA, FALSE,
  "performance_vs_market", "Performance versus market", 1, NA, FALSE,
  "off_epa_per_play", "Offensive EPA per play", 1, "offensive_plays", TRUE,
  "def_epa_allowed", "Defensive EPA allowed per play", -1, "defensive_plays", TRUE,
  "off_success_rate", "Offensive success rate", 1, "offensive_plays", TRUE,
  "def_success_rate_allowed", "Defensive success rate allowed", -1, "defensive_plays", TRUE,
  "cpoe", "CPOE", 1, "pass_attempts", FALSE,
  "off_sack_rate", "Sack rate allowed", -1, "offensive_dropbacks", FALSE,
  "def_sack_rate", "Defensive sack rate", 1, "defensive_dropbacks", FALSE
)
```

```{r}
#| label: outcome-guide
#| code-fold: true

df_metric_spec |>
  transmute(
    Outcome = label,
    `Season-average weight` = coalesce(weight, "Equal games"),
    `Better performance` = if_else(direction == 1, "Higher", "Lower"),
    Standardized = if_else(standardize, "Yes", "No")
  ) |>
  knitr::kable()
```

## Eligible plays

All play-based measures use regular-season plays with both teams recorded, excluding quarterback kneels and spikes. Additional requirements depend on the measure:

- **EPA and success rate:** run or pass plays with EPA recorded. Eligible overtime, two-point, and penalty plays are retained.
- **CPOE:** pass attempts with completion status and CPOE recorded.
- **Sack rates:** quarterback dropbacks, including scrambles, with sack status recorded.

Defensive measures summarize the opposing offense’s performance. Defensive CPOE is calculated for opponent adjustment but is not reported as a separate outcome.

The common eligibility filters are applied once per season and reused below.

```{r}
#| label: prepare-regular-plays

pbp_regular_list <- map(pbp_list, function(df_pbp) {
  df_reg_plays <- df_pbp |>
    filter(
      season_type == "REG",
      !coalesce(qb_kneel == 1, FALSE),
      !coalesce(qb_spike == 1, FALSE),
      !is.na(posteam), posteam != "",
      !is.na(defteam), defteam != ""
    )
  return(df_reg_plays)
})
```

## Calculate game-level performance

Metrics are calculated for every team in the reference seasons because opponent adjustment requires league-wide results. For 2026, calculations are limited to Melbourne and the two following games.

```{r}
#| label: calculate-observed-metrics

reference_seasons <- unique(df_reference_team_seasons$season)
df_league_game_metrics <- map_dfr(reference_seasons, function(year) {
  df_reg_plays <- pbp_regular_list[[as.character(year)]]
  calculate_game_metrics(df_reg_plays, df_team_schedules)
}) 

df_historical_registry <- df_game_registry |>
  filter(season < 2026) |>
  select(season, team, game_id, phase)

df_melbourne_registry <- df_game_registry |>
  filter(season == 2026) |>
  select(season, team, game_id, phase)

df_melbourne_observed <- pbp_regular_list[["2026"]] |>
  semi_join(df_melbourne_registry, by = "game_id") |>
  calculate_game_metrics(df_team_schedules)
```

# Opponent adjustments

The seven play-based outcomes are adjusted for the opponent’s performance relative to the league:

$$\text{Adjusted value} = \text{Observed value}
- (\text{Opponent average} - \text{League average}).$$

Offensive outcomes use the opponent’s defensive average, and defensive outcomes use its offensive average. Scoring margin and market residual are not adjusted.

```{r}
#| label: opponent-specification

df_opponent_spec <- tribble(
  ~metric, ~opponent_metric, ~opponent_weight, ~league_metric,
  "off_epa_per_play", "def_epa_allowed", "defensive_plays", "off_epa_per_play",
  "def_epa_allowed", "off_epa_per_play", "offensive_plays", "off_epa_per_play",
  "off_success_rate", "def_success_rate_allowed", "defensive_plays", "off_success_rate",
  "def_success_rate_allowed", "off_success_rate", "offensive_plays", "off_success_rate",
  "cpoe", "def_cpoe_allowed", "defensive_pass_attempts", "cpoe",
  "off_sack_rate", "def_sack_rate", "defensive_dropbacks", "off_sack_rate",
  "def_sack_rate", "off_sack_rate", "offensive_dropbacks", "off_sack_rate"
)

df_opponent_spec <- df_opponent_spec |> 
  left_join(
    df_metric_spec |> select(league_metric = metric, league_weight = weight),
    by = "league_metric",
    relationship = "many-to-one"
  )
```

For historical and reference-season games, the opponent average excludes the game being assessed; the league average includes the full season. Melbourne and its following games use complete 2025 opponent and league averages.

Averages are weighted by eligible plays, pass attempts, or dropbacks, as appropriate for each measure.

```{r}
#| label: calculate-adjusted-metrics

df_team_game_metrics <- df_league_game_metrics |>
  semi_join(df_reference_team_seasons, by = join_by(season, team)) |>
  adjust_for_opponent(df_league_game_metrics) |>
  arrange(season, team, gameday)

df_historical_game_metrics <- df_historical_registry |>
  left_join(
    df_team_game_metrics,
    by = join_by(season, team, game_id),
    relationship = "one-to-one"
  )

df_melbourne_game_metrics <- df_melbourne_registry |>
  left_join(
    df_melbourne_observed,
    by = join_by(season, team, game_id),
    relationship = "one-to-one"
  ) |>
  adjust_for_opponent(df_league_game_metrics, reference_season = melbourne_reference_season)
```

# Season comparisons

Each selected game is compared with the same team’s reference season. Unadjusted outcomes use unadjusted reference values, and adjusted outcomes use adjusted reference values.

| Comparison | Weighting | Reference games |
|------------------------|------------------------|------------------------|
| Season averages in original units | Equal games for scoring margin and market residual; opportunity-weighted for play-based outcomes | Full reference season |
| Standardized differences and counts of less favorable or tied games | Equal games | Excludes the focal game for historical comparisons; all 17 games from 2025 for Melbourne |
| Historical scoring and market averages in Figure 3 | Equal games | Excludes both the London game and its first following game |

EPA and success-rate outcomes are standardized using the reference mean and sample standard deviation:

$$\text{Standardized difference} =
\text{Direction}\times
\frac{\text{Game value}-\text{Reference mean}}
{\text{Reference standard deviation}}.$$

Direction is +1 when higher values are favorable and −1 when lower values are favorable. Positive standardized differences therefore always indicate better performance. CPOE and sack rates are reported without standardization.

The following chunks specify the unadjusted and adjusted outcomes, calculate season averages and standardized differences, and count reference games with less favorable or equal performance. Figure 3’s season averages are calculated separately using the exclusions above.

```{r}
df_comparison_spec <- bind_rows(
  df_metric_spec |> mutate(column = metric, adjustment = "Unadjusted"),
  df_metric_spec |>
    semi_join(df_opponent_spec, by = "metric") |>
    mutate(column = paste0("adj_", metric), adjustment = "Opponent adjusted")
) |> 
  arrange(match(metric, df_metric_spec$metric), adjustment == "Opponent adjusted")
```

```{r}
#| label: calculate-comparisons

df_melbourne_comparisons <- compare_with_reference(
  df_melbourne_game_metrics, 
  df_team_game_metrics
)

df_historical_comparisons <- compare_with_reference(
  df_historical_game_metrics, 
  df_team_game_metrics
)

df_historical_score_context <- df_team_game_metrics |>
  semi_join(df_historical_registry, by = join_by(season, team)) |>
  anti_join(df_historical_registry, by = join_by(season, team, game_id)) |>
  summarise(
    across(c(scoring_margin, performance_vs_market), mean),
    n_games = n(),
    .by = c(season, team)
  )
```

# Results

The tables report observed values, season averages, and comparisons with individual reference games. `games_exceeded` counts games with less favorable performance; `games_tied` counts equal values. Success and sack rates are displayed as percentages.

```{r}
#| label: results-display
#| code-fold: true

format_outcome <- function(metric, value) {
  formatted_values <- case_when(
    str_detect(metric, "success|sack") ~ sprintf("%.1f%%", 100 * value),
    metric == "cpoe" ~ sprintf("%+.1f pp", value),
    str_detect(metric, "epa") ~ sprintf("%+.3f", value),
    .default = sprintf("%+.1f", value)
  )
  return(formatted_values)
}

show_comparisons <- function(df_comparisons) {
  df_display <- df_comparisons |>
    transmute(
      season, team, phase, outcome = label, adjustment,
      observed = format_outcome(metric, target_value),
      season_mean = format_outcome(metric, season_mean),
      reference_n_games, games_exceeded, games_tied,
      standardized_difference = round(standardized_score, 2)
    )
  comparison_table <- rmarkdown::paged_table(df_display)
  return(comparison_table)
}

offensive_outcomes <- c("off_epa_per_play", "off_success_rate", "cpoe", "off_sack_rate")
defensive_outcomes <- c("def_epa_allowed", "def_success_rate_allowed", "def_sack_rate")
df_melbourne_matchup <- df_melbourne_comparisons |> filter(phase == "international")
```

## Melbourne episode

### Game summary

The tables show each team’s score, closing expected margin, and market residual, followed by comparisons with its 2025 season.

```{r}
#| label: melbourne-game-summary
#| code-fold: true

df_melbourne_game_metrics |>
  filter(phase == "international") |>
  select(
    team,
    opponent,
    points_for,
    points_against,
    expected_margin,
    scoring_margin,
    performance_vs_market
  ) |> 
  knitr::kable()
```

```{r}
#| code-fold: true

df_melbourne_matchup |>
  filter(metric %in% c("scoring_margin", "performance_vs_market")) |>
  show_comparisons()
```

### 49ers offense versus Rams defense

The 49ers’ offensive EPA and success rate describe the same plays as the Rams’ defensive values. Each team is compared with its own reference season, with and without opponent adjustment. CPOE and sack rates are also shown.

```{r}
#| label: san-francisco-offense
#| code-fold: true

df_melbourne_matchup |>
  filter(
    (team == "SF" & metric %in% offensive_outcomes) |
      (team == "LA" & metric %in% defensive_outcomes)
  ) |>
  show_comparisons()
```

### Rams offense versus 49ers defense

The corresponding comparisons are shown for the Rams’ offense and the 49ers’ defense.

```{r}
#| label: los-angeles-offense
#| code-fold: true

df_melbourne_matchup |>
  filter(
    (team == "LA" & metric %in% offensive_outcomes) |
      (team == "SF" & metric %in% defensive_outcomes)
  ) |>
  show_comparisons()
```

### Play and sack counts

The table shows the plays, pass attempts, dropbacks, and sacks contributing to each team’s Melbourne results.

```{r}
#| code-fold: true

df_melbourne_game_metrics |>
  filter(phase == "international") |>
  select(
    team,
    offensive_plays,
    pass_attempts,
    offensive_dropbacks,
    sacks_allowed,
    defensive_plays,
    defensive_dropbacks,
    sacks_generated
  ) |> 
  knitr::kable()
```

### Melbourne performance (Figure 2)

Figure 2 combines the outcome relative to the closing spread with EPA and success rates in original and standardized units. Positive standardized values indicate better performance.

```{r}
#| label: melbourne-performance
#| fig-width: 9
#| fig-height: 9.4
#| code-fold: true

melbourne_figure <- build_melbourne_performance_figure(
  df_melbourne_matchup |> filter(adjustment == "Unadjusted"),
  df_melbourne_matchup |> filter(adjustment == "Opponent adjusted"),
  df_melbourne_game_metrics |> filter(phase == "international")
)

save_paper_figure(
  melbourne_figure,
  "melbourne_outcome_performance",
  width = 9,
  height = 9.4,
  directory = figure_directory
)

print(melbourne_figure)
```

### First games after Melbourne

Each team’s first following game uses the same 2025 reference season and opponent-adjustment approach as the Melbourne game.

```{r}
#| label: post-melbourne-results
#| code-fold: true

df_melbourne_game_metrics |>
  filter(phase == "return") |>
  select(team, opponent, gameday, points_for, points_against, performance_vs_market) |>
  knitr::kable()
```

```{r}
#| code-fold: true

df_melbourne_comparisons |>
  filter(phase == "return") |>
  show_comparisons()
```

## Historical episodes

### Travel context (Table 1)

Table 1 contains the manually compiled travel context and reported results for the five London episodes. Each trip followed a road game, and each London game was followed by a bye week.

```{r}
#| code-fold: true

df_historical_context_data <- tibble::tribble(
  ~team_year, ~opponent_location, ~staging_context, ~arrival_nights,
  ~travel_approach, ~pregame_context, ~international_result,
  ~post_rest, ~first_post_result,
  "49ers, 2010", "Denver Broncos; Wembley Stadium, London",
  "Played at Carolina Oct 24; flew directly from Charlotte after the game",
  "Mon Oct 25; time not reported; 6 nights",
  "Early arrival; spent game week in London",
  "49ers 1–6; Broncos 2–5",
  "Won 24–16; +6.0 points",
  "Week 9 bye; 14 days between games",
  "Won 23–20 (OT) vs St Louis; −1.5 points",
  "49ers, 2013", "Jacksonville Jaguars; Wembley Stadium, London",
  "Played at Tennessee Oct 20; direct overnight flight from Nashville",
  "Mon Oct 21, about 13:00; 6 nights",
  "Early arrival; spent game week in London",
  "49ers 5–2; Jaguars 0–7",
  "Won 42–10; +16.0 points",
  "Week 9 bye; 14 days between games",
  "Lost 9–10 vs Carolina; −7.0 points",
  "Rams, 2017", "Arizona Cardinals; Twickenham Stadium, London",
  "Played at Jacksonville Oct 15; stayed there through Thursday",
  "Fri Oct 20, about 09:00; 2 nights",
  "Brief London stay; prepared in eastern US on usual weekly schedule",
  "Rams 4–2; Cardinals 3–3",
  "Won 33–0; +30.0 points",
  "Week 8 bye; 14 days between games",
  "Won 51–17 at New York Giants; +28.5 points",
  "Rams, 2019", "Cincinnati Bengals; Wembley Stadium, London",
  "Played at Atlanta Oct 20; stayed through Thursday practice",
  "Fri Oct 25 morning; exact time not reported; 2 nights",
  "Brief London stay; same eastern-US preparation pattern as 2017",
  "Rams 4–3; Bengals 0–7",
  "Won 24–10; +2.0 points",
  "Week 9 bye; 14 days between games",
  "Lost 12–17 at Pittsburgh; −9.0 points",
  "Rams, 2025", "Jacksonville Jaguars; Wembley Stadium, London",
  "Played at Baltimore Oct 12; stayed and practiced there during game week",
  "<30 hours before kickoff; exact time not reported; 1 night",
  "Brief London stay; prepared in eastern US, aiming to maintain usual routine",
  "Rams 4–2; Jaguars 4–2",
  "Won 35–7; +25.0 points",
  "Week 8 bye; 14 days between games",
  "Won 34–10 vs New Orleans; +10.0 points"
)
```

```{r}
#| label: historical-context-table
#| code-fold: true

(gt_context_table <- historical_context_table(df_historical_context_data))
```

### Early-arriving 49ers

The 2010 and 2013 London games and their first following games are compared with the 49ers’ performance in the corresponding seasons.

```{r}
#| label: historical-san-francisco
#| code-fold: true

df_historical_comparisons |>
  filter(team == "SF") |>
  show_comparisons()
```

### Short-stay Rams

The same comparisons are shown for the Rams’ 2017, 2019, and 2025 episodes.

```{r}
#| label: historical-los-angeles
#| code-fold: true

df_historical_comparisons |>
  filter(team == "LA") |>
  show_comparisons()
```

### Scoring margin and market residual (Figure 3)

Figure 3 compares the London and following games with season averages excluding both games. The accompanying table shows those averages and the number of reference games.

```{r}
#| label: historical-score-context
#| code-fold: true

df_historical_score_context |> knitr::kable(digits = 2)
```

```{r}
#| label: historical-outcomes
#| fig-width: 9
#| fig-height: 5.85
#| code-fold: true

historical_figures_list <- build_historical_figures(
  df_historical_game_metrics,
  df_historical_score_context,
  df_historical_comparisons
)

save_paper_figure(
  historical_figures_list$outcomes, 
  "historical_score_market_baselines", 
  width = 9, 
  height = 5.85, 
  directory = figure_directory
)

save_paper_figure(
  historical_figures_list$performance,
  "historical_international_return_play_metrics",
  width = 9, 
  height = 6.25, 
  directory = figure_directory
)

print(historical_figures_list$outcomes)
```

### Standardized performance (Figure 4)

Figure 4 shows offensive and defensive EPA and success rates relative to the other games in each team’s season, with and without opponent adjustment.

```{r}
#| label: historical-performance
#| fig-width: 9
#| fig-height: 6.25
#| code-fold: true

historical_figures_list$performance
```

# Exports

Input information and result bundles are saved as RDS files; source information and analysis tables are saved as CSV files. Archive outputs go to `data/derived/`. Fresh nflverse outputs go to `outputs/nflverse/<timestamp>/data/derived/`.

```{r}
#| label: export-analysis
#| code-fold: true

inputs_list <- list(
  snapshot_date = snapshot_date,
  provenance = df_data_source,
  pbp_paths = pbp_paths,
  schedule_path = schedule_path,
  schedules = df_schedules,
  game_registry = df_game_registry,
  reference_team_seasons = df_reference_team_seasons
)

historical_results_list <- list(
  snapshot_date = snapshot_date,
  provenance = df_data_source,
  league_game_metrics = df_league_game_metrics,
  team_game_metrics = df_team_game_metrics,
  game_metrics = df_historical_game_metrics,
  comparisons = df_historical_comparisons,
  context = df_historical_score_context
)

melbourne_results_list <- list(
  snapshot_date = snapshot_date,
  provenance = df_data_source,
  game_metrics = df_melbourne_game_metrics,
  comparisons = df_melbourne_comparisons
)

analysis_exports_list <- list(
  historical_game_metrics = df_historical_game_metrics,
  historical_reference_comparisons = df_historical_comparisons,
  historical_score_market_context = df_historical_score_context,
  melbourne_game_metrics = df_melbourne_game_metrics,
  melbourne_reference_comparisons = df_melbourne_comparisons
)

dir.create(derived_directory, recursive = TRUE, showWarnings = FALSE)
write_csv(df_data_source, file.path(derived_directory, "data_source.csv"), na = "")
saveRDS(inputs_list, file.path(derived_directory, "validated_inputs.rds"))
saveRDS(historical_results_list, file.path(derived_directory, "historical_results.rds"))
saveRDS(melbourne_results_list, file.path(derived_directory, "melbourne_results.rds"))

iwalk(analysis_exports_list, function(df_data, name) {
  path <- file.path(derived_directory, paste0(name, ".csv"))
  write_csv(df_data, path, na = "")
})
```

The following chunk exports Table 1 as PNG and Word files. Statistical figures were saved as PNGs in their results sections. Fresh nflverse runs save these exports within their timestamped run folder. The Melbourne travel figure is maintained separately.

```{r}
#| label: export-context-table
#| code-fold: true

dir.create(table_directory, recursive = TRUE, showWarnings = FALSE)
gt::gtsave(
  gt_context_table,
  file.path(table_directory, "table_1_historical_international_context.png"),
  vwidth = 1300, 
  vheight = 900, 
  expand = 10
)
save_historical_context_docx(
  gt_context_table,
  file.path(table_directory, "table_1_historical_international_context.docx")
)
```
